---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpbayes

Fast rejection sampling and Bayesian inference for COM-Poisson regression models [Benson and Friel (2021)](https://projecteuclid.org/journals/bayesian-analysis/volume-16/issue-3/Bayesian-Inference-Model-Selection-and-Likelihood-Estimation-using-Fast-Rejection/10.1214/20-BA1230.full). Regression structure on the location and dispersion parameters is supported and MCMC sampling relies on the exchange algorithm.

Please ensure proper setup of a C++ compiler as our package links to Rcpp and RcppArmadillo (you might test this by installing these two first). Uncomment the following command line to install `cpbayes` from this GitHub repository.

```{r setup}
#devtools::install_github("luizapiancastelli/cpbayes")
library(cpbayes)

ls("package:cpbayes") #List of currently exported functions
```

Usage of `rcompois` for set parameter values (the fast-rejection sampler):

```{r}
Y = rcompois(n =100, mu =1, nu = 0.5)
```

### COM-Poisson regression

Function `fitcpbayes` fits a COM-Poisson GLM and a quick check of its arguments is rendered by `help("fitcpbayes")` (documentation under development!). The code below provides and example of its usage with the `takeoverbids` data set (model 4 in (Benson and Friel (2021)) ) and some methods currently available.

```{r}
#help("fitcpbayes")
data("takeoverbids")

mcmc = fitcpbayes(numbids ~ whtknght, numbids ~ size, takeoverbids, 10000, 10000, nchains =3)

summary(mcmc) #Gives the number of efficient draws neff and scale reduction factors Rhat
```
```{r}
plot(mcmc)
plot(mcmc, type = 'trace')
```

```{r}
BIC(mcmc)
```

