---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cpbayes

Fast rejection sampling and Bayesian inference for COM-Poisson regression models [Benson and Friel (2021)](https://projecteuclid.org/journals/bayesian-analysis/volume-16/issue-3/Bayesian-Inference-Model-Selection-and-Likelihood-Estimation-using-Fast-Rejection/10.1214/20-BA1230.full). Regression structure on the location and dispersion parameters is supported and MCMC sampling relies on the exchange algorithm.

Please ensure proper setup of a C++ compiler as our package links to Rcpp and RcppArmadillo (you might test this by installing these two first). Uncomment the following command line to install `cpbayes` from this GitHub repository.

```{r setup}
#devtools::install_github("luizapiancastelli/cpbayes")
library(cpbayes)
ls("package:cpbayes") #List of currently exported functions
```

### COM-Poisson distribution

`cpbayes` provides a family of functions designed to work with the COM-Poisson distribution, namely `rcompois`, `dcompois`, `pcompois` and `qcompois`. A call to `help` renders documentation where further details on their implementation can be found. 

`rcompois` is the key component among these, an implementation of the fast-rejection sampler of Benson and Friel (2021) that gives \bold{exact draws} from a COM-Poisson with location $\mu$ and dispersion parameter $\nu$. Currently `dcompois`, `pcompois` and `qcompois` involve approximation of the normalisation constant via truncation, included so that `cpbayes` gives a complete framework to work with the COM-Poisson.

The next code chunk exemplifies simulating over and underdispersed counts with `rcompois` and computation of some COM-Poisson probabilities.

```{r}
Y1 = rcompois(n =200, mu =1, nu = 0.2) #nu < 1 renders overdispersed count data
Y2 = rcompois(n =200, mu =1, nu = 1.5) #nu > 1 underdispersion
c(mean(Y1), var(Y1))
c(mean(Y2), var(Y2))
#approximate mean and variance are E(Y) ~ mu + 1/2nu -1/2 and Var(Y) ~ mu/nu

#P(Y = 0)
dcompois(0, mu =1, nu = 0.2)
dcompois(0, mu =1, nu = 1.5)

#P(Y <= 1) for Y1, Y2
pcompois(1, mu = 1, nu= 0.2)
pcompois(1, mu = 1, nu=1.5)

#Median of distribution 
qcompois(0.5, 1, 0.2)
qcompois(0.5, 1, 1.5)
```

### COM-Poisson regression

In addition to the COM-Poisson distribution functions, `cpbayes` fits generalised linear models to COM-Poisson distributed counts in the Bayesian framework. Regression structure via the logarithm link is supported in the location and dispersion parameters. This model assumes counts $y_1, \cdots, y_n$ to be COM-Poisson distributed with likelihood

$$
    f(y_i|\mu_i, \nu_i) =  \left( \dfrac{\mu_i^{y_i}}{y_i!} \right)^{\nu_i} \frac{1}{Z(\mu_i, \nu_i)}
$$

where $Z(\mu_i, \nu_i) = \sum_{y=0}^\infty q(y_i|\mu_i, \nu_i)$, $q(y_i|\mu_i, \nu_i) =  (\frac{\mu_i^{y_i}}{y_i!})^{\nu_i}$.

Regression structure on the location is $\mu_i = \exp ( \beta_{\mu,0} + \sum_{j=1}^b \beta_{\mu, j} x^\mu_{ij} )$ and similarly $\nu_i = \exp ( \beta_{\nu,0} + \sum_{k=1}^c \beta_{\nu, k} x^\nu_{ik} )$ for the dispersion. Here $\boldsymbol{x}^\mu_1, \cdots, \boldsymbol{x}^\mu_b$ and $\boldsymbol{x}^\nu_1, \cdots, \boldsymbol{x}^\nu_c$ are observed covariates to include in $\mu$ and $\nu$ for explaining $\boldsymbol{y}$.

The function `fitcpbayes` runs Markov Chain Monte Carlo for $\boldsymbol{\beta_\mu}, \boldsymbol{\beta_\nu}$ Bayesian inference using the exchange algorithm (Murray (2006)). Further details on the exchange framework for intractable likelihoods and arguments to `fitcpbayes` are rendered by `help("fitcpbayes")`.

Code below provides and example of its usage with the `takeoverbids` data, a data set on the number of bids `numbids` received by U.S. firms in the period 1978-1985 that were taken over within 1 year of the initial offer. This is modeled alongside the potential covariates `whtknght` and `size` described in `help(takeoverbids)`. 

`fitcpbayes`'s basic arguments include `R` formulas for $\mu$ and $\nu$, the data set (a `data.frame`), number of iterations to discard as `burnin` and to store (`niter`). `nchains` specifies how many MCMC chains to run and parallel computation can be enabled by setting `ncores`. `fitcpbayes` outputs an object of class `cpbayes` for which generic `summary` and `plot` methods are available.

```{r}
data("takeoverbids")
mcmc = fitcpbayes(numbids ~ whtknght, numbids ~ size, takeoverbids, 10000, 10000, nchains =3)

summary(mcmc) #neff is the number of efficient draws, and Rhat the potential scale reduction factor, estimated if nchains>1
```

```{r}
plot(mcmc) #density plots using combined posterior draws
plot(mcmc, type = 'trace') #trace plots colored by chain run
```

`cpbayes` also provides BIC estimation for comparing between models fits. This is automatic from a `cpbayes` object, where likelihood estimation occurs is linked to the acceptance rate of COM-Poisson rejection sampling. For more details, please see the indicated reference or `help("BIC")`. Let's consider another regression structure for the `takeoverbids` problem and use this tool for model selection.  

```{r}
mcmc2 = fitcpbayes(numbids ~ whtknght + bidprem, numbids ~ size, takeoverbids, 10000, 10000, nchains =3)

BIC(mcmc)
BIC(mcmc2)
```

It looks like the data further supports `numbids ~ whtknght` $\mu$ regression in comparison to `numbids ~ whtknght + bidprem`. 

#### No regression

A special case of `fitcpbayes` is implemented for when we are interested in Bayesian inference for the  $\mu, \nu$ parameters with no regression structure. As an example, let's revisit `Y2` and see how the model fit supports the true values used to generate the data, these were $\mu =1, \nu = 1.5$.

```{r}
y2 = data.frame('y' = Y2)
mcmc3 = fitcpbayes(y ~ 1, y ~ 1, y2, 10000, 10000, nchains =1)

summary(mcmc3)
```
Looks like the true values have high posterior probability as expected!
